# Reproducible Research Project1
## Setting the environments
```{r}
#setwd("D:/My Documents/GitHub/GitHub/Reproducible Research/Course Project 1")
library(ggplot2)
Sys.setlocale(category = "LC_ALL", locale = "english")
```

## Loading and preprocessing the data
```{r}
#fileurl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#download.file(fileurl, "./activity.zip", mode = "wb")
#unzip("activity.zip")
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date)
```

## What is mean total number of steps taken per day?
### Calculating total number of steps taken per day
```{r}
dailySteps <- tapply(activity$steps, activity$date, FUN = sum, na.rm = TRUE)
qplot(dailySteps, binwidth = 1000, xlab = "total number of steps taken each day")
```

### Mean and Median of daily steps
```{r}
mean(dailySteps)
median(dailySteps)
```

## Calculating the average daily activity pattern
```{r}
averSteps <- aggregate(x = list(steps = activity$steps), by = list(interval = activity$interval), FUN = mean, na.rm = TRUE)
p <- ggplot(averSteps, aes(x = interval, y = steps))
p + geom_line() + xlab("5-minut interval") + ylab("average number of steps taken")
maxInterval <- averSteps[which.max(averSteps$steps), ]
print(maxInterval)
```

## Inputting missing values and generate new dataset
```{r}
missingValue <- is.na(activity$steps)
summary(missingValue)

filledValue <- function(steps, interval)
{
    #filled <- NA
    if(!is.na(steps))
      filled <-c(steps)
    else filled <- (averSteps[averSteps$interval == interval, "steps"])
    return(filled)
}

filledActivity <- activity
filledActivity$steps <- mapply(filledValue, filledActivity$steps, filledActivity$interval)
filledDailySteps <- tapply(filledActivity$steps, filledActivity$date, FUN = sum)
qplot(filledDailySteps, binwidth = 1000, xlab = "total number of filled steps per day")
```

## Comparing the differences in activity pattern between weekday and weekend
```{r}
weekend <- weekdays(as.Date(filledActivity$date)) %in% c("Saturday", "Sunday")
filledActivity$daytype <- "weekday"
filledActivity$daytype[weekend ==TRUE] <- "weekend"
filledActivity$daytype <- as.factor(filledActivity$daytype)

weekdayCheck <- function(date)
{
    day  <- weekdays(date)
    if(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
      return("weekday")
    else if (day %in% c("Saturday", "Sunday"))
      return("weekend")
      else stop("invalid date")
}

filledActivity$date <- as.Date(filledActivity$date)
filledActivity$day <- sapply(filledActivity$date, FUN = weekdayCheck)
```

## Final Plot with the difference between weekdays and weekends
```{r}
diff <- aggregate(steps ~ interval + day, data = filledActivity, mean)
p <- ggplot(diff, aes(interval, steps))
p + geom_line() + facet_grid(day ~ .) + xlab("5-minute interval") + ylab("number of steps")
```

